Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-06-12 paper/paper.tex
Nuke old plots

\begin{figure}[t]
    \includegraphics*[scale=0.7]{plots/33v-par-cumulative.pdf}
    \hfill
    \includegraphics*[scale=0.7]{plots/plain-par-cumulative.pdf}

    \vspace*{1em}
    \includegraphics*[scale=0.7]{plots/33ved-par-cumulative.pdf}
    \hfill
    \includegraphics*[scale=0.7]{plots/plain-connected-par-cumulative.pdf}

    \vspace*{1em}
    \includegraphics*[scale=0.7]{plots/33ve-connected-par-cumulative.pdf}
    \hfill
    \includegraphics*[scale=0.7]{plots/sip-par-cumulative.pdf}

    \caption{The cumulative number of instances solved in under a certain time, comparing sequential
    (solid lines) and parallel (dotted lines) algorithms, for six different families of maximum
    common (connected) subgraph problems. All experiments use 32 threads on a 16 core,
    hyper-threaded system.}
    \label{figure:mcs-par-unconnected-cumulative}
\end{figure}

Cumulative performance plots comparing each parallel algorithm to its sequential version are given
in \cref{figure:mcs-par-unconnected-cumulative}, using 32 threads on 16 core hyper-threaded Intel Xeon
E5-2640 v2 systems. The results are consistent: in each case the parallel algorithm is clearly and
substantially better than its sequential version, except when the sequential runtime is below 100
milliseconds.

\begin{figure}[p]
    \centering
    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/plain-par-heatmap-clique.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-kdown.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-james.pdf}\hspace*{0.5em}

    \vspace*{2em}
    \centering
    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/33ved-par-heatmap-clique.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/33ved-par-heatmap-james.pdf}\hspace*{0.5em}

    \vspace*{2em}
    \centering
    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/sip-par-heatmap-kdown.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/sip-par-heatmap-james.pdf}\hspace*{0.5em}

    \caption{The benefits of parallelism, for different sets of problem instances and for different
    algorithms. All experiments use 32 threads on a 16 core, hyper-threaded system; points below the
    top diagonal line indicate a speedup, and the lower diagonal line is a sixteen times speedup.}
    \label{figure:mcs-par-speedups}
\end{figure}

\Cref{figure:mcs-par-speedups} gives instance-by-instance comparisons for a representative subset of
these algorithm and problem combinations. For the clique algorithm on unlabelled graphs (top left),
the results show fairly consistent speedups for harder instances: there is only a single instance
whose sequential runtime is over two seconds where we obtain a speedup of less than ten. We also see
a significant number of superlinear speedups. This is because there are relatively long proofs of
optimality for most of the harder instances, but strong solutions are usually found quickly (and
where they are not, the additional diversity helps). The results are similarly positive on
labelled graphs (centre left), with speedups of greater than ten being the norm for harder
instances, although there is more variety in the results.

@@

\begin{figure}[p]
    \centering
    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-clique-par-repeat-heatmap.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/plain-clique-par-8to32-heatmap.pdf}\hspace*{1em}

    \vspace*{1em}
    \centering
    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-kdown-par-repeat-heatmap.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/plain-kdown-par-8to32-heatmap.pdf}\hspace*{1em}

    \vspace*{1em}
    \centering
    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-james-par-repeat-heatmap.pdf}\hfill
    \includegraphics*[scale=0.7]{plots/plain-james-par-8to32-heatmap.pdf}\hspace*{1em}

    \caption{Parallel search is reproducible and scalable, as well as risk-free and beneficial. In the
    left-hand column, we compare each of the three parallel algorithms against a repeated run with
    the same parameters, on the unconnected, unlabelled instances (points on the diagonal line show
    reproducibility of runtimes). In the right-hand column, we compare going from 8 threads to 32
    threads (points below the diagonal line are an improvement). All experiments are on a 16 core,
    hyper-threaded system.}
    \label{figure:mcs-par-repeats}
\end{figure}

@@

\Cref{figure:mcs-par-unconnected-cumulative,figure:mcs-par-speedups} show that for all three
algorithms, parallel search is beneficial, and also (reasonably) risk-free: although in a few cases the
constant factor slowdowns due to hardware limitations and having to make changes to data structures
are unpleasant, we do not encounter any exponential slowdowns due to search tree changes. But what
about reproducibility and scalability? Further experiments in \cref{figure:mcs-par-repeats}
establish that both of these properties also hold. In the left-hand column, we plot each algorithm
run against itself, using 32 threads in both cases. For the clique and $k{\downarrow}$ approaches,
every instance is on the central diagonal line, showing an extremely high level of reproducibility.
For the splitting algorithm, there is a little more variability, particularly for easier instances:
inspecting the results suggests that this is not down to different amounts of search work being
performed, but rather whether or not helper threads end up triggering additional copying.

In the right-hand column of \cref{figure:mcs-par-repeats}, we plot the effects of moving from 8
threads (on the same system) to 32 threads. Points below the diagonal line indicate an improvement.
For the clique approach, increasing the number of threads is unequivocally beneficial. For the
$k{\downarrow}$ algorithm, the benefits are slim, and sometimes performance worsens slightly due to
memory contention (not work done) but we do not see any exponential slowdowns. Finally, for the
splitting algorithm, the results are usually strictly better, although in a few cases the overheads
increase further giving a slowdown; we also see an increase in the occurrence of strongly
superlinear speedups.

>>>>

Cumulative performance plots comparing each parallel algorithm to its sequential version are given
in ??, using 32 threads on 16 core hyper-threaded Intel Xeon
E5-2640 v2 systems. The results are consistent: in each case the parallel algorithm is clearly and
substantially better than its sequential version, except when the sequential runtime is below 100
milliseconds.

?? gives instance-by-instance comparisons for a representative subset of
these algorithm and problem combinations. For the clique algorithm on unlabelled graphs (top left),
the results show fairly consistent speedups for harder instances: there is only a single instance
whose sequential runtime is over two seconds where we obtain a speedup of less than ten. We also see
a significant number of superlinear speedups. This is because there are relatively long proofs of
optimality for most of the harder instances, but strong solutions are usually found quickly (and
where they are not, the additional diversity helps). The results are similarly positive on
labelled graphs (centre left), with speedups of greater than ten being the norm for harder
instances, although there is more variety in the results.

@@

@@

?? show that for all three
algorithms, parallel search is beneficial, and also (reasonably) risk-free: although in a few cases the
constant factor slowdowns due to hardware limitations and having to make changes to data structures
are unpleasant, we do not encounter any exponential slowdowns due to search tree changes. But what
about reproducibility and scalability? Further experiments in ??
establish that both of these properties also hold. In the left-hand column, we plot each algorithm
run against itself, using 32 threads in both cases. For the clique and $k{\downarrow}$ approaches,
every instance is on the central diagonal line, showing an extremely high level of reproducibility.
For the splitting algorithm, there is a little more variability, particularly for easier instances:
inspecting the results suggests that this is not down to different amounts of search work being
performed, but rather whether or not helper threads end up triggering additional copying.

In the right-hand column of ??, we plot the effects of moving from 8
threads (on the same system) to 32 threads. Points below the diagonal line indicate an improvement.
For the clique approach, increasing the number of threads is unequivocally beneficial. For the
$k{\downarrow}$ algorithm, the benefits are slim, and sometimes performance worsens slightly due to
memory contention (not work done) but we do not see any exponential slowdowns. Finally, for the
splitting algorithm, the results are usually strictly better, although in a few cases the overheads
increase further giving a slowdown; we also see an increase in the occurrence of strongly
superlinear speedups.
